---
title: "General Social Survey"
author: "Corie Christianson"
date: "Year 2019"
output: 
  html_document:
    number_sections: true
    toc: true
---


```{r include=FALSE,echo=FALSE}
require(tigerstats)
require(tidyverse)
require(foreign)
Gss <- read.spss("GSS2018.sav")
Gss1 <- as.tibble(Gss)
```

```{r include=FALSE,echo=FALSE}
#names(Gss1)
mydata <- select(Gss1,VOTE16,RACE,REGION,EDUC,BIGBANG,EVOLVED,DEGREE,POSTLIFE,AGE,OWNGUN,CAPPUN,COURTS,POLVIEWS,VOTE12,GOD)
mydata$AGE <- as.numeric(mydata$AGE)
mydata$EDUC <- as.numeric(mydata$EDUC)
mydata2 <-filter(mydata,RACE!="IAP" ) %>% droplevels()
mydata3 <- filter(mydata2,BIGBANG=="True" | BIGBANG=="False",EVOLVED=="True"|EVOLVED=="False",POSTLIFE=="YES"|POSTLIFE=="NO",OWNGUN=="YES"|OWNGUN=="NO",COURTS=="TOO HARSH"|COURTS=="NOT HARSH ENOUGH"|COURTS=="ABOUT RIGHT") %>% droplevels()
mydata4 <- filter(mydata3,CAPPUN=="FAVOR"|CAPPUN=="OPPOSE",VOTE12=="Voted"|VOTE12=="Did not vote",VOTE16=="Voted"|VOTE16=="Did not vote") %>% droplevels()
levels(mydata4$VOTE12)[1] <- "voted12"
levels(mydata4$VOTE12)[2] <- "no in 12"

```


```{r}
barchartGC(~VOTE16+VOTE12, data=mydata4, type = "percent") 
chisqtestGC(~VOTE16+VOTE12, data=mydata4)
votetab=xtabs(~VOTE16+VOTE12, data=mydata4)
fisher.test(votetab)
```

# Introduction Question 1

<!--In this section you explain what you are trying to show.  Where did the data come from?  What is the research or other question you are trying to answer?!-->

This data is from the 2018 General Social Survey. The first blocks of R-code has selected down a subset of the data to just 15 variables. It has further removed unwanted factor levels in much of the data. Examine that code to see what it is doing. First explain what you make out of comparing how people voted in 2012 with 2016.


## Methods

<!--Decide on your methods:  use "variable analysis" or other appropriate descriptors.  Make sure to choose at least one graphical method and at least one numerical method.!-->

The methods used in this experiment were Pearson's Chi-squared test with Yates' continuity correction,  Fisher's Exact Test for Count Data, and barcharts for graphical descriptive results

##Results

<!--Divide this section into two sub-sections:  One for your descriptive  results and one for your inferential results.!-->

### Descriptive Results
There was an obvious relationship between those that voted in 2012 were most assuredly going to vote again in 2016.

#### Graphical Descriptive Results

<!--Graphical results here.  Make sure to show your code.  Provide appropriate labels for axes, giving units if possible, and provide a good title for the graph, too.  Use the graphical results to describe the patterns if any that exist in the data as focused toward the research question!-->

```{r}
voted1=c(92.04,35.96) 
voted2=c(35.96,14.04)
tabl = rbind(voted1,voted2)
barchartGC(tabl)
```

#### Numerical Descriptive Results

<!--Numerical results go here. Use the numerical results to describe the patterns if any that exist in the data as focused toward the research question!-->

We get row and column percents
```{r}
mytab <- xtabs(~VOTE16+VOTE12, data=mydata4)
rowPerc(mytab)
colPerc(mytab)
```


### Inferential Results

<!--State hypothesis clearly.  Make sure your discussion of the inferential test covers all the aspects that the test output produces, such as test statistic, p-value etc.  Make a decision about the null hypothesis, explain the assumptions on which the selected test/procedure was based, and why the chosen procedure satisfys the assumptions and is appropriate to answer the research question!-->
We assume the null hypothesis in this case

$H_0$ : There is a relationship between those that voted in 2012 and those that voted in 2016.

$H_A$ : There is no relationship between those that voted in 2012 and 2016.
```{r}
chisqtestGC(~VOTE16+VOTE12, data=mydata4)
fisher.test(mytab)
```

# Introduction Question 2

<!--In this section you explain what you are trying to show.  Where did the data come from?  What is the research or other question you are trying to answer?!-->

Now answer Is belief in the bigbang theory (BIGBANG) independent of belief in the afterlife (POSTLIFE)

## Methods

<!--Decide on your methods:  use "variable analysis" or other appropriate descriptors.  Make sure to choose at least one graphical method and at least one numerical method.!-->

The methods used in this experiment were Pearson's Chi-squared test with Yates' continuity correction,  Fisher's Exact Test for Count Data, and barcharts for graphical descriptive results

##Results

<!--Divide this section into two sub-sections:  One for your descriptive  results and one for your inferential results.!-->

### Descriptive Results

There is no clear relationship that those whoe believe in the Big Bang Theory also believe in Post Life. The two beliefs are independent in nature. 

#### Graphical Descriptive Results

<!--Graphical results here.  Make sure to show your code.  Provide appropriate labels for axes, giving units if possible, and provide a good title for the graph, too.  Use the graphical results to describe the patterns if any that exist in the data as focused toward the research question!-->
```{r}
barchartGC(~BIGBANG+POSTLIFE, data=mydata4, type = "percent") 
barchartGC(~BIGBANG+POSTLIFE, data=mydata4)
```

#### Numerical Descriptive Results

<!--Numerical results go here. Use the numerical results to describe the patterns if any that exist in the data as focused toward the research question!-->

We get row and column percents
```{r}
mytab <- xtabs(~BIGBANG+POSTLIFE, data=mydata4)
rowPerc(mytab)
colPerc(mytab)
```
### Inferential Results

<!--State hypothesis clearly.  Make sure your discussion of the inferential test covers all the aspects that the test output produces, such as test statistic, p-value etc.  Make a decision about the null hypothesis, explain the assumptions on which the selected test/procedure was based, and why the chosen procedure satisfys the assumptions and is appropriate to answer the research question!-->

We assume the null hypothesis in this case

$H_0$ : There is no relationship between believing in the BIGBANG Theory and beliving in POSTLIFE.

$H_A$ : There is a relationship between believing in the BIGBANG Theory and beliving in POSTLIFE.
```{r}
chisqtestGC(~BIGBANG+POSTLIFE, data=mydata4)
fisher.test(mytab)
```

# Introduction Question 3

<!--In this section you explain what you are trying to show.  Where did the data come from?  What is the research or other question you are trying to answer?!-->

This time compare the opinion on death penalty (CAPPUN) with opinion on courts (COURTS).

## Methods

<!--Decide on your methods:  use "variable analysis" or other appropriate descriptors.  Make sure to choose at least one graphical method and at least one numerical method.!-->

The methods used in this experiment were Pearson's Chi-squared test with Yates' continuity correction,  Fisher's Exact Test for Count Data, and barcharts for graphical descriptive results.

##Results

<!--Divide this section into two sub-sections:  One for your descriptive  results and one for your inferential results.!-->

### Descriptive Results
There is no clear relationship that those whoe own a gun also believe in capitol punishment. The two beliefs are independent in nature.

#### Graphical Descriptive Results

<!--Graphical results here.  Make sure to show your code.  Provide appropriate labels for axes, giving units if possible, and provide a good title for the graph, too.  Use the graphical results to describe the patterns if any that exist in the data as focused toward the research question!-->
```{r}
barchartGC(~OWNGUN+CAPPUN, data=mydata4, type = "percent") 
barchartGC(~OWNGUN+CAPPUN, data=mydata4)
```

#### Numerical Descriptive Results

<!--Numerical results go here. Use the numerical results to describe the patterns if any that exist in the data as focused toward the research question!-->

We get row and column percents
```{r}
mytab <- xtabs(~OWNGUN+CAPPUN, data=mydata4)
rowPerc(mytab)
colPerc(mytab)
```
### Inferential Results

<!--State hypothesis clearly.  Make sure your discussion of the inferential test covers all the aspects that the test output produces, such as test statistic, p-value etc.  Make a decision about the null hypothesis, explain the assumptions on which the selected test/procedure was based, and why the chosen procedure satisfys the assumptions and is appropriate to answer the research question!-->

We assume the null hypothesis in this case

$H_0$ : There is no relationship between gun ownership and belief in capital punishment.

$H_A$ : There is a relationship between gun ownership and belief in capitol punishment.

```{r}
chisqtestGC(~OWNGUN+CAPPUN, data=mydata4)
fisher.test(mytab)
```
For the fisher exact test we have 

$H_0$ : The odds of owning a gun and favoring capitol punishment are one to one. 

$H_A$ : The odds of owning a gun and favoring capitol punishment do not contain the number one.

```{r}
fisher.test(mytab)
```


# Discussion and Conclusion

<!--Discussion and conclusion here.  If you found a relationship be sure to consider whether the relationship occurs because one of the variavbles causes the other, or whether they perhasps are related for some other reason.  Watch the chapter 6 videos from the GeorgeTown videos collection.!-->

For the first question, I found that there was an obvious relationship between those that voted in 2012 were most assuredly going to vote again in 2016.For the second questions, I found that there was an obvious relationship between those that voted in 2012 were most assuredly going to vote again in 2016.For the last question, I found that there is no clear relationship that those whoe own a gun also believe in capitol punishment. The two beliefs are independent in nature.